---
title: "Exercise 12-1"
author: "Gao"
date: "2023-10-06"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, we must install our libraries

```{r}
library(tidyverse)
library(tidymodels)
```

Set working directory

```{r}
getwd()
```

Retrieve the .csv file

```{r}
housing <- read_csv("melbourne_housing.csv")

names(housing) <- names(housing) %>% str_to_title()
```

Examine for missing values

```{r}
apply(X = housing, MARGIN = 2, FUN = function(col) 
  sum(is.na(col)))
```

Drop all rows with missing values

```{r}
housing <- na.omit(housing)
```

Return correlation 

```{r}
cor(x = housing$Price, y = housing$Buildingarea)
cor(x = housing$Price, y = housing$Rooms)
cor(x = housing$Price, y = housing$Car)
cor(x = housing$Price, y = housing$Landsize)
```

Display a scatter pot that shows the relationshiop between the Price and BuildingArea variables

```{r}
ggplot(housing, aes(x=Price, y=Buildingarea)) + geom_point()
```

Split the data

```{r}
housing_split <- initial_split(housing, prop = 0.75)
train <- training(housing_split)
test <- testing(housing_split)
```

Create functions to delete outliers

```{r}
get_upper_fence <- function(x) {
  quantile(x, 0.75) + (1.5 * IQR(x))
}

get_lower_fence <- function(x) {
  quantile(x, 0.25) - (1.5 * IQR(x))
}
```

Use the fences to delete outliers

```{r}

train <- train %>% 
  filter_at(vars(Price, Buildingarea), 
            all_vars(. > get_lower_fence(.) &
                       . < get_upper_fence(.)))
test <- test %>% 
  filter_at(vars(Price, Buildingarea), 
            all_vars(. > get_lower_fence(.) &
                       . < get_upper_fence(.)))
```

Delete any really weird numbers

```{r}
train <- train %>% filter(Buildingarea > Price * (1/20000) & Price < Buildingarea * 100000)
test <- test %>% filter(Buildingarea > Price * (1/20000) & Price < Buildingarea * 100000)

ggplot(train, aes(y=Price, x=Buildingarea)) + geom_point()
```

Create a linear regression model using the training set

```{r}
linear_reg()


model <- fit(object = linear_reg(), formula = Price ~ Buildingarea, 
             data = train)
predict(model, new_data = test)
```

Use the testing set

```{r}
model_results <- test %>% 
  mutate(predict(model, new_data = test))
```

Plot the plot

```{r}
ggplot(data = model_results) + 
  geom_point(aes(x = Buildingarea, y = Price)) + 
  geom_point(aes(x = Buildingarea, y = .pred), color = "blue")
```

