---
title: "Medical Malpractice"
author: "Gao"
date: "2023-10-09"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import libraries

```{r cars}
library(tidyverse)
library(tidymodels)
```

Import the data

```{r pressure, echo=FALSE}
getwd()

malpractice <- read_csv("malpractice.csv") %>% as_tibble()
```

Create a plot

```{r}
ggplot(malpractice) + geom_point(aes(x = Severity, y = Amount))
```

Create a linear regression model

```{r}
model <- lm(Amount ~ Severity, data = malpractice)
model
```

points-hat = 7471 + 31255(GP)

Interpretation of the slope:

For each additional point increase in severity, we expect that the malpractice claim amount increases by 31255 dollars. 

Interpretation of the y-intercept

When a case has a severity of 0, the expected payout is 7471 dollars. This has no practical interpretations.

```{r}
cor(malpractice$Severity, malpractice$Amount)
```

r = 0.3371629

There is a moderately strong positive linear correlation between games played by an NBA player and average points per game.

```{r}
(cor(malpractice$Severity, malpractice$Amount))^2
```

r^2 = 0.1136788

Approximately 11.37% of the variability in the mean points per game can be explained by this lienar model containing games played and mean points per game.

```{r}
summary(model)
```

New plot containing 

```{r}
ggplot(malpractice) + geom_point(aes(x = Severity, y = resid(model)))
```


