---
title: "NBA"
author: "Gao"
date: "2023-10-09"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import libraries

```{r cars}
library(tidyverse)
library(tidymodels)
```

Import the data

```{r pressure, echo=FALSE}
getwd()

NBA <- read_csv("NBA.csv") %>% as_tibble()
```

Create a plot

```{r}
ggplot(NBA) + geom_point(aes(x = GP, y = PTS))
```

Create a linear regression model

```{r}
model <- lm(PTS ~ GP, data = NBA)
model
```

points-hat = 2.253 + 0.153(GP)

Interpretation of the slope:

For each additional game played for an NBA player, we expect that his average points per game to increase by 0.153 points. 

Interpretation of the y-intercept

When a player doesn't play any games, he will have an average point per game of 2.2528. This has no practical interpretations.

```{r}
cor(NBA$GP, NBA$PTS)
```

r = 0.544

There is a moderately strong positive linear correlation between games played by an NBA player and average points per game.

```{r}
(cor(NBA$GP, NBA$PTS))^2
```

r^2 = 0.2955

Approximately 29.55% of the variability in the mean points per game can be explained by this lienar model containing games played and mean points per game.

```{r}
summary(model)
```

New plot containing 

```{r}
ggplot(NBA) + geom_point(aes(x = GP, y = resid(model)))
```

Interpret residual plot

