---
title: "Rollercoaster"
author: "Gao"
date: "2023-10-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import libraries or something

```{r}
library(tidyverse)
library(tidymodels)
library(ggforce)
library(yardstick)
library(car)
library(moments)
library(GGally)
library(psych)
library(fastDummies)
```

Import the data

```{r}
rollercoasters <- read_csv("rollercoasters.csv") %>% as_tibble()
```
Find correlation between Speed and all numerical values

```{r}
cor(rollercoasters$Speed, select_if(rollercoasters, is.numeric))
```

Find the regression model

```{r}
model <- lm(Speed ~ Height + Drop + Length + Duration + Inversions, data = rollercoasters)
summary(model)
```

Construct the residual plot

```{r}
plot(fitted(model), resid(model))
```
Draw outlier boundaries

```{r}
standard_error <- sqrt(deviance(model)/df.residual(model))
standard_error
2*standard_error
plot(fitted(model),resid(model))
abline(h=2*standard_error, col = "red")
abline(h=-2*standard_error, col = "red")
```

2x standard error of mean: 8.068208
There are 5 outliers that are outside of 2 standard errors of the mean. (Coasters 19, 2, 16, 43, 3)
Display residual values

```{r}
residuals <- resid(model)
sort(residuals)
```

Cook's Distance

```{r}

```


Jackknife

```{r}
jac <- qt(df= 101 - 5 - 2, 0.95)
jac
sort(jackknife <- rstudent(model))
```

Jackknife = 1.661226
Outliers: 19, 2, 16, 17, 24, 95, 53, 54, 43, 3

Leverage

```{r}
lev <- 2 * (1+5)/101
lev
sort(hatvalues(model))
```
leverage = 0.1188119
Outliers: 3, 98, 7, 99, 1, 93, 92, 94

QQPlot

```{r}
qqnorm(resid(model))
```

```{r}

```

```{r}
model <- lm(Speed ~ Height + Drop + Length + Duration + Inversions, data = rollercoasters)
model
```

Speed-hat = 0.0407 * Height + 0.14685 * Drop + 0.00131 * Length + -0.2937 * Duration + -0.28856 * Inversions + 35.79567

```{r}
rollercoasters <- dummy_cols(rollercoasters, select_columns = "Track", remove_first_dummy = TRUE)
rollercoasters
```

```{r}
model2 <- lm(Speed ~ Height + Drop + Length + Duration + Inversions + Track_Wood, data = rollercoasters)
summary(model2)
model2 %>% tidy()
```

Multiple R-Squared Adjusted: 0.9183
F-Value: 118.5
P-Value: 2.2 * 10 ^ -16

These values suggest that the model has a very strong fit

```{r}
plot(fitted(model), resid(model))
```

```{r}
Houses <- read_csv("houses.csv") %>% as_tibble()
summary(Houses)
```

