---
title: "Airlines"
author: "Gao"
date: "2024-02-02"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(fastDummies)
```

```{r}
Airlines <- read_csv("Airline.csv") %>% as_tibble()
summary(Airlines)
print(Airlines)
```
```{r}
Airlines <- dummy_cols(Airlines, select_columns = "satisfaction", remove_first_dummy = TRUE)

Airlines <- dummy_cols(Airlines, select_columns = "Gender", remove_first_dummy = TRUE)
Airlines <- dummy_cols(Airlines, select_columns = "CustomerType", remove_first_dummy = TRUE)
Airlines <- dummy_cols(Airlines, select_columns = "TypeofTravel", remove_first_dummy = TRUE)
Airlines <- dummy_cols(Airlines, select_columns = "Class", remove_first_dummy = TRUE)

Airlines
```


\newpage

1. Fit your model.

```{r}
model1 <- glm(satisfaction_satisfied ~ Age + Gender_Male + FlightDistance + ArrivalDelayinMinutes + Class_Eco + Class_EcoPlus + CustomerType_LoyalCustomer + TypeofTravel_PersonalTravel, family = "binomial", data = Airlines)
summary(model1)
model1
```

$\operatorname{logit}\left[ \widehat{\text{pr}}\left( Y=1 \right) \right]=0.4083332-1.0953468\left( \text{Gender} \right)-0.0029852\left( \text{Age} \right)-0.0050386\left( \text{ArrivalDelay} \right)-0.0001284\left( \text{FlightDistance} \right)-1.3029748\left( \text{ClassEco} \right)-1.4153322\left( \text{ClassEcoPlus} \right)+1.9241037\left( \text{LoyalCustomer} \right)-0.2753098\left( \text{PersonalTravel} \right)$

\newpage

2. Compute and interpret the estimate odds ratio for an airline passenger’s satisfaction who are male and those who are female, controlling for age, flight distance, arrival delay and the other independent variables that you added to your model.

```{r}
OR <- model1$coefficients[["Gender_Male"]]
print(exp(OR))
```

The estimated odds of a male being satisfied with a flight is 0.3344 times that of a female.

\newpage

3. Construct and interpret a 98% confidence interval for the estimate odds ratio in question 2.

```{r}

beta1 <- model1$coefficients[["Gender_Male"]]
print(beta1)

zcrit <- qnorm(0.99)
print(zcrit)

stddevbeta1 <- coef(summary(model1))[2, "Std. Error"]
print(stddevbeta1)

lb <- exp(beta1 - zcrit*stddevbeta1)
ub <- exp(beta1 + zcrit*stddevbeta1)

print(lb)
print(ub)

print(paste0("The 98% confidence interval for the estimate odds ratio in part (B) is (", round(lb, 4), ", ", round(ub, 4), ")"))
```

This means that we are 98% confident that the odds ratio lies between 0.3341 and 0.3348. Since 1 is not included in this range, we can say that we are 98% confident that being male plays a factor in customer ratings.

\newpage

4. Test the null hypothesis that the adjusted odds ratio from question 2 equals 1. Use the Wald test.

```{r}

library(lmtest)

waldtest(model1, test="Chisq", terms=2)

```

$\text{H}_{0}: \beta_{1} = 0$

$\text{H}_{a}: \beta_{1} \ne 0$

$\text{T.S.: } \chi^2 = 6882.7$

$\text{p-value: } P(\chi^2 \ge 1.196) \approx 2.2 * 10^{-16}$

Conclusion: Since the p-value < $\alpha$, we reject $\text{H}_{0}$. At the $\alpha$ = 0.05 significance level there is sufficient evidence to indicate that being male plays a factor in customer satisfaction

\newpage

5. Test the null hypothesis that the adjusted odds ratio from question 2 equals 1. Use the LR test.

```{r}

model1 <- glm(satisfaction_satisfied ~ Age + Gender_Male + FlightDistance + ArrivalDelayinMinutes + Class_Eco + Class_EcoPlus + CustomerType_LoyalCustomer + TypeofTravel_PersonalTravel, family = "binomial", data = Airlines)

summary(model1)

model2 <- glm(satisfaction_satisfied ~ Age + FlightDistance + ArrivalDelayinMinutes + Class_Eco + Class_EcoPlus + CustomerType_LoyalCustomer + TypeofTravel_PersonalTravel, family = "binomial", data = Airlines)

summary(model2)

152934-145667

1-pchisq(7267,df=1)
```

$\text{H}_{0}: \beta_{1} = 0$

$\text{H}_{a}: \beta_{1} \ne 0$

$\text{T.S.: } \text{LR}=-2\log {{\hat{L}}_{R}}-\left( -2\log {{{\hat{L}}}_{C}} \right)=152934-145667=7267$

$\text{p-value: } P(\chi^2 \ge 1.21) \approx 0$

Conclusion: Since the p-value <<< $\alpha$, we reject $\text{H}_{0}$. At the $\alpha$ = 0.05 significance level there is sufficient evidence to indicate that being male plays a factor in customer satisfaction

\newpage

6. Calculate and interpret the estimated odds ratio for two people, one with an age of 18 and the other a 51 years old, controlling for gender, flight distance, arrival delay and the other independent variables you added to your model.

```{r}
OR2 <- model1$coefficients[["Age"]]
OR2
print(exp(33*OR2))
```

${{\widehat{OR}}_{\left( \text{AGE=51 }-\text{ AGE=18 = 33  }\text{  | } \text{  GEN, SAL} \right)}}={{e}^{33{{\beta }_{2}}}}={{e}^{33\left( -0.0030 \right)}}\approx 0.906$

Someone who is 51 years old is 0.906 times as likely to be satisfied with a flight than someone who is 18 years old.

\newpage

7. Fit a logistic model that regresses satisfaction on the predictors gender, age, arrival delay and the interaction term between gender and age.

```{r}
model3 <- glm(satisfaction_satisfied ~ Age + Gender_Male + ArrivalDelayinMinutes + Gender_Male:Age, family = "binomial", data = Airlines)
summary(model3)
model3
```
$\operatorname{logit}\left[ \widehat{\text{pr}}\left( Y=1 \right) \right]=0.070659-0.937770\left( \text{Gender} \right)-0.016246\left( \text{Age} \right)-0.004808\left( \text{ArrivalDelay} \right)+0.001237\left( \text{Gender:Age} \right)$

8. Using the model from question 7, compute and interpret the 92% confidence interval for the estimated adjusted odds ratio for the effect of gender, controlling for age and arrival delay when a person is 40 years old.


```{r}

vcov(model3)

qnorm(0.96)

```

$\exp \left[ {{{\hat{\beta }}}_{1}}+{{{\hat{\beta }}}_{4}}\left( \text{Gender} \right)\pm 1.751\left( {{\left( \text{Gender} \right)}^{2}}\cdot \widehat{\text{Var}}\left( {{{\hat{\beta }}}_{1}} \right)+{{\left( \text{Age} \right)}^{2}}\cdot \widehat{\text{Var}}\left( {{{\hat{\beta }}}_{4}} \right)+2\left( \text{Gender} \right)\left( \text{Age} \right)\cdot \widehat{\text{Cov}}\left( {{{\hat{\beta }}}_{1}},{{{\hat{\beta }}}_{4}} \right) \right) \right]$

$=\exp \left[ -0.937770+0.001237\left( 40 \right)\pm 1.751\left( {{\left( 1 \right)}^{2}}\cdot {{\left( 3.046951*10^{-7} \right)}}+{{\left( 40 \right)}^{2}}\cdot {{\left( 5.970795*10^{-7} \right)}}+2\left( 1 \right)\left( 40 \right)\left( -3.046848*10^{-7} \right) \right) \right]$

$=\exp \left[ -0.88829\pm 0.00163063 \right]$

$=\left( {{e}^{-0.88829-0.00163063}},{{e}^{-0.88829+0.00163063}} \right)$

$=\left( 0.411,0.412 \right)$