---
title: "SUV"
author: "Worcester"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tidymodels)
library(ggforce)
library(mctest)
library(olsrr)
library(jtools)
library(ggcorrplot)          
library(yardstick)
library(car)
library(moments)
library(GGally)
library(psych)
library(fastDummies)
library(lmtest)
```

```{r}
SUV <- read_csv("SUV.csv") %>% as_tibble()
summary(SUV)
SUV
```

#Question 1

#Can you predict whether or not a person purchases a SUV (1 = purchased a SUV, 0 = did not purchase a SUV) based on their age (AGE), annual salary (SAL) and gender (GEN) (Male = 1, Female = 0).

#Part A

#Fit the logistic regression model that regresses SUV on the predictors AGE, SAL and GEN.

```{r}
model1 <- glm(SUV ~ GEN + SAL + AGE, family = "binomial",  data = SUV)
summary(model1)
model1
```

SUV = -1.278 + 3.38 * 10 ^ -1 * GENERATION + 3.644 * 10 ^ -5 * SALARY + 2.370 * 10 ^ -1 * AGE

#Part B

#Compute the estimate odds ratio for purchasing a SUV for persons who are male and those who are female, controlling for age and salary.

```{r}
exp(3.338*10^-1)
```

#Part C

#Interpret your answer in part (B) in the context of the problem.

It is the odds ratio of a dude getting a SUV versus a girl getting a SUV. Someone who is Male is 1.4 times more likely to get an SUV than someone who is Female.

#Part D

#Construct a 95% confidence interval for the estimate odds ratio in part (B).

```{r}
qnorm(0.975)
exp(3.338*10^-1 + 1.959964 * 3.052*10^-1)
exp(3.338*10^-1 - 1.959964 * 3.052*10^-1)
```

(0.767681, 2.539535)

There is a 95% certainty that the odds ratio between a guy getting an SUV and a girl getting a SUV lies between 0.767681 and 2.539535

#Part E

#Test the null hypothesis that the adjusted odds ratio from part (B) equals 1.  Use the Wald test.

The null hypothesis is that there is no statistically significant difference between the chance of a guy buying a SUV and a girl buying a SUV.

```{r}
waldtest(model1, test="Chisq", terms = 1)
```

Since the p-value, 0.2741, exceeds the alpha value of 0.05, we fail to reject the HO!!! Our confidence interval has insufficient evidence to determine that there is a difference in the probability of purchasing an SUV depending on if the buyer is a male of female.

#Part F

#Test the null hypothesis that the adjusted odds ratio from part (B) equals 1.  Use the LR test.

```{r}
model2 <- glm(SUV ~ SAL + AGE, family = "binomial",  data = SUV)
summary(model2)
model2

277.05-275.84

1-pchisq(1.21, df=1)

```

Since the p-value, 1.21, exceeds the alpha value, we once again fail to reject the HO!!! Our confidence interval has insufficient evidence to determine that there is a difference in the probability of purchasing an SUV depending on if the buyer is a male of female.

#Part G

#Calculate the estimated odds ratio for two people, one with an age of 40 and the other a 52 year old, controlling for salary and gender.

```{r}
exp(12*2.370*10^-1)
```

#Part H

#Fit a logistic model that regresses SUV on the predictors AGE, SAL, GEN and the interaction term between AGE and GEN.

```{r}
model3 <- glm(SUV ~ GEN + SAL + AGE + AGE:GEN, family = "binomial",  data = SUV)
summary(model3)
model3
```

SUV = -12.13 - 1.214 * Generation + 3.657 * 10 ^ -5 * Salary + 2.21 * 10 ^ -1 * Age + 3.835 * 10 ^ -2 * Gen:Age

#Part I

#Using the model from (H), compute the 95% confidence interval for the estimated adjusted odds ratio for the effect of gender, controlling for AGE and SAL when a person is 25 years old.

```{r}
round(vcov(model3), 3)

1.877^2 + 25^2*(4.593 * 10 ^ -2) + 2 * 1.877 * (-0.085)

exp(-1.214 + 25*3.84*10^-2 + 1.96 * (1.877^2 + 25^2 * 0.0459^2 + 2*25*(-0.0851)))
exp(-1.214 + 25*3.84*10^-2 - 1.96 * (1.877^2 + 25^2 * 0.0459^2 + 2*25*(-0.0851)))
```

We are 95% confident that the actual odds ratio lies between 0.2465045 and 2.44092. 

#Part J

#What does the confidence interval in part (I) tell you?

There is a 95% certainty that the odds ratio lies within the interval. Since the interval includes 1, it means that there is a chance that there is no difference in SUV purchasing between genders when all other variables are controlled for.