---
title: "Healthcare Analysis"
author: "Gao"
date: "2024-02-28"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidymodels)
library(tidyverse)
library(rpart.plot)
library(caret)
library(dplyr)
```

1-5.

```{r}
Watson <- read_csv("Watson_Healthcare_Modified.csv")
```

6. Convert the Attrition column from the character type to the factor type. This is necessary because the dependent variable in a classification model must be of the factor type.

```{r}
Watson$Attrition <- as.factor(Watson$Attrition)
```

7. Drop all columns of the character type that only have a single unique value. This is necessary because having an independent variable that only has a single unique values causes an error if you attempt to tune a classification model.

I chose these three variables by looking at the tibble in the environment pane

```{r}
Watson <- select(Watson,-c(EmployeeCount, Over18, StandardHours))
```

8. Split the data into testing and training sets

I used a 75-25 split

```{r}
Watson_split <- initial_split(Watson, prop = 0.75)

train <- training(Watson_split)

test <- testing(Watson_split)
```

9. Create a decision tree model with a depth of 30 and a minimum of 5 data points per node

```{r}
dtree_mod <- decision_tree(mode = "classification", min_n = 5, tree_depth = 30)
```

10. Fit the model to predict the Attrition variable given all of the remaining variables in the data set.

```{r}
dtree_mod_fit <- fit(object = dtree_mod, formula = Attrition ~ ., data = train)
```

\newpage

11. Plot the model.

```{r}
dtree_mod_fit %>% extract_fit_engine() %>% rpart.plot()
```

\newpage

12. Create the predictions for the model.

```{r}
predictions <- predict(dtree_mod_fit, new_data = test) 

predictions
```

\newpage

13. View the confusion matrix for the model as well as the accuracy and kappa metrics.

```{r}
confusionMatrix(data = predictions$.pred_class, reference = test$Attrition)
```

\newpage

14. Check the variable importance of the model. Make a note of any variables that have an importance of 0.

As we can see, Education, PerformanceRating, RelationshipSatisfaction and TrainingTimesLastYear all have an importance of 0

```{r}
dtree_mod_fit %>% extract_fit_engine() %>% varImp()
```

15. Create six cross validation folds for the training data

```{r}
folds <- vfold_cv(train, v = 6)
```

\newpage

16. Create an automatic grid of hyperparameters for the model

```{r}
tree_grid_auto <- grid_regular(tree_depth(), min_n(), cost_complexity())

tree_grid_auto

print(tree_grid_auto, n = 27)
```

\newpage

17. Use the grid and the folds to tune the model

```{r}
dtree_mod_tune <- decision_tree(mode = "classification", tree_depth = tune(), min_n = tune(), cost_complexity = tune())

tuning_data <- tune_grid(dtree_mod_tune, preprocessor = Attrition ~., resamples = folds, grid = tree_grid_auto)

print(tuning_data %>% collect_metrics, n = 54)
```

18. Display the best hyperparameter values for the model

The best hyperparameter values are min_n = 40, and tree_depth = 8

```{r}
tuning_data %>% select_best(metric = "accuracy")

tuning_data %>% show_best(metric = "accuracy")
```

19. Create the final model that uses the best hyperparameter values.

```{r}
dtree_mod_final <- decision_tree(mode = "classification", min_n = 40, tree_depth = 8)
```


20. Fit the final model without using any variables that have an importance of 0.

```{r}
dtree_mod_fit_final <- fit(object = dtree_mod, formula = Attrition ~ . - Education - PerformanceRating - RelationshipSatisfaction - TrainingTimesLastYear, data = train)
```

\newpage

21. Use the final mode to create predictions.

```{r}
predictions_final <- predict(dtree_mod_fit_final, new_data = test) 

predictions_final
```

22. View the confusion matrix for the final model

```{r}
confusionMatrix(data = predictions_final$.pred_class, reference = test$Attrition)
```

